% chapter 6
\chapter{Evaluations and Conclusions}
\section{A brief Summary}
In this project, we aim to analyse a given VRP problem using chosen LP tools and analyse them in terms of their performance
and features. We have set some goals to fulfill those aims and the methodology to achieve those goals.

By following the methodology, we have successfully created LP formulations and models using
chosen tools. Our models have meet all of the stipulated requirements that were set in chapter 3.
In performance analysis, we have found that Gurobi and Optaplanner give optimal solution to solve the given VRP instances. On the other hand,
or-tools consistently underperform on most of the datasets, especially the P-VRP-60. In addition, Gurobi was impractical to use
as it takes a very long time to compute the solution. Given these reasons, we have decided to use Optaplanner to solve the given
VRP problem. During the implentation of our LP models, we have also conducted feature analysis and tabulated the results in table
5.4.

Using optaplanner, we found the optimal solution to the problem to be 10.27 for CVRP and 10.49 for CVRPTW model measured in eucledian distances.
The real distances in km are 798.57 and 828.95 for CVRP and CVRPTW model respectively. Unfortunately, those distances does not
accurately represent the actual road distances as we assume the distance from one vertex to another is straight. This assumption
was made due to lack of traffic data. In addition to the total distance, we have also
tabulated the set of routes and route visualisation for each model.

\section{Evaluation}
In the list below, we asses the outcome of this project based on the goals stated in chapter 1.
\begin{enumerate}
    \item \textbf{Goal 1 and 2} - We have reviewed the literature and the tools that are needed to understand linear programming and create an
     formulation to the given VRP problem. We can confirm this by the fact that our LP models have produced solutions that
     are very close to the best.
    \item \textbf{Goal 3} - We have conducted performance analysis and feature analysis on all the chosen LP tools. The results
    are shown in chapter 5.
    \item \textbf{Goal 4} - We have successfully produced the CVRP and CVRPTW models using Optaplanner that produces the
     total minimal distance travel and the optimal vehicle routes based on the given VRP problem.
\end{enumerate}

\section{Future Work}
There are clearly more room for improvements in some aspects of this project, which would be ideal for future work.
The list of possible future works are listed below:
\begin{enumerate}
    \item Using real road distances or other cost functions such as fuel or time taken in the VRP model.
    \item Solving VRP by using other methods such as dynamic or genetic programming.
    \item Studying the effects of heuristics on the solution of VRP or other routing problems.
\end{enumerate}